---
title: "Modeling Comorb"
author: "Team 6"
date: '2022-07-15'
output: html_document
---

```{r}
setwd("C:/Users/lizvi/Cal State Fullerton/Jaynes, Jessica - CSUF CHOC Summer 2022 Research/Group2")

load("cfCOVIDgroup6.RDATA")

library(tidyverse)
library(viridis)
library(glmnet)
library(gee)
library(caret)
library(InformationValue)
```


```{r message = FALSE}
attach(d)

all_comorb <- d %>% 
  select(encounterid, personid, servicedate, contains("comorb"),
         COVIDseverity, age_at_encounter) %>% 
  mutate(servicedate = as.Date(d[,"servicedate"]), 
         COVIDseverity = as.ordered(case_when(COVIDseverity > 0 ~ 1,
                                   COVIDseverity == 0 ~ 0)),
         personid = as.factor(personid),
         age_group = as.factor(case_when(age_at_encounter <= 11 ~ "0 - 11",
                               age_at_encounter > 11 & 
                                 age_at_encounter <= 17 ~ "12 - 17",
                               age_at_encounter > 17 &
                                 age_at_encounter < 21 ~ "18 - 20",
                               age_at_encounter > 20 ~ "21 - 25")),
         payer = as.factor(payer))

total_comorb_df <- dplyr::select(d, contains("comorb")) %>% 
  mutate_if(is.factor, as.numeric) %>% 
  - 1
  
total_comorb_df <- total_comorb_df %>% 
  mutate(total_comorb_by_patient = c(rowSums(total_comorb_df)), 
         encounterid = d[,"encounterid"]) %>% 
  dplyr::select("encounterid", "total_comorb_by_patient")

first_enc <- all_comorb %>%
  arrange_at("servicedate") %>%
  distinct(personid, .keep_all = TRUE) %>% 
  left_join(total_comorb_df, by = "encounterid") %>% 
  mutate(total_comorb_by_patient = as.factor(
    case_when(total_comorb_by_patient == 0 ~ "0",
              total_comorb_by_patient == 1 ~ "1",
              total_comorb_by_patient == 2 ~ "2",
              total_comorb_by_patient > 2 ~ "3+")))

table(first_enc$age_at_encounter)/length(first_enc$age_at_encounter)

cumsum(table(first_enc$age_at_encounter)/length(first_enc$age_at_encounter))  

first_enc_all <- d %>% 
   mutate(servicedate = as.Date(d[,"servicedate"]), 
          personid = as.factor(personid)) %>% 
  arrange_at("servicedate") %>% 
  distinct(personid, .keep_all = TRUE)
          
```


Shows total occurrence of each comorbidity in data set
```{r warning = FALSE}
totalco <- first_enc[,4:24] %>% 
  mutate_if(is.factor, as.numeric) %>%
  - 1 %>%
  colSums()

totalco
```


```{r}
table(first_enc$payer)

glm(COVIDseverity ~ payer, data = first_enc, family = "binomial") %>% 
  summary()
```


```{r}
glm(COVIDseverity ~., data = first_enc[,c(4:26, 28)], family = "binomial") %>% 
  summary()
```


Logistic Regression now with grouped ages based on proportion. Each age group is
roughly 25% of the sample. 
```{r}

glm(COVIDseverity ~ ., 
               data = first_enc[,c(4:25, 27:28)], family = "binomial") %>% 
  summary()
```


With ages grouped by WHO standard
```{r}
all_comorb_WHO <- d %>% 
  select(encounterid, personid, servicedate, contains("comorb"),
         COVIDseverity, age_at_encounter) %>% 
  mutate(servicedate = as.Date(d[,"servicedate"]), 
         COVIDseverity = as.ordered(case_when(COVIDseverity > 0 ~ 1,
                                   COVIDseverity == 0 ~ 0)),
         personid = as.factor(personid),
         age_group = as.factor(case_when(age_at_encounter <= 4 ~ "0 - 4",
                               age_at_encounter > 4 & 
                                 age_at_encounter < 10 ~ "5 - 9",
                               age_at_encounter > 9 &
                                 age_at_encounter < 15 ~ "10 - 14",
                               age_at_encounter > 14 &
                                 age_at_encounter < 20 ~ "15 - 19",
                               age_at_encounter > 19 ~ "20 - 25")),
         payer = as.factor(payer))

first_enc_WHO <- all_comorb_WHO %>%
  distinct(personid, .keep_all = TRUE)

glm(COVIDseverity ~ ., 
               data = first_enc_WHO[,c(4:25, 27:28)], family = "binomial") %>% 
  summary()
```


LASSO ?!!!?!!!
```{r}
set.seed(622)
x <-  model.matrix(COVIDseverity ~., first_enc[,c(4:28)])[, -1]
y <-  first_enc$COVIDseverity

grid <- 10^seq(10, -2, length = 100)

train <-  sample(nrow(x), size = floor(.8 * nrow(x)))
test <-  (-train)
y_test <-  y[test]

lasso_mod <-  glmnet(x[train, ], y[train], alpha = 1,
                     lambda = grid, family = "binomial")

cv_out <- cv.glmnet(x[train, ], y[train], alpha = 1, family = "binomial")

best_lam <-  cv_out$lambda.min

lasso_pred <-  predict(lasso_mod, s = best_lam, newx = x[test, ])

out <-  glmnet(x, y, alpha = 1, lambda = grid, family = "binomial")

lasso_coef = predict(out, type = "coefficients",
                         s = best_lam)[1:27, ]

lasso_coef[lasso_coef != 0]
```

```{r}


olr_df <- d %>% 
  select(contains("comorb"),
         COVIDseverity, age_at_encounter) %>% 
  mutate(servicedate = as.Date(d[,"servicedate"]),
         personid = as.factor(personid),
         age_group = as.factor(case_when(age_at_encounter <= 11 ~ "0 - 11",
                               age_at_encounter > 11 & 
                                 age_at_encounter <= 17 ~ "12 - 17",
                               age_at_encounter > 17 &
                                 age_at_encounter < 21 ~ "18 - 20",
                               age_at_encounter > 20 ~ "21 - 25")),
         payer = as.factor(payer)) %>% 
  arrange_at("servicedate") %>%
  distinct(personid, .keep_all = TRUE)

library(MASS)

train <-  sample(nrow(olr_df), size = floor(.80 * nrow(olr_df)))

dtrain <- olr_df[train,]
test <- olr_df[-train,]

olr_fit <- polr(COVIDseverity ~ comorb_bronchiectasis_J47 +
                comorb_resp_failure_J96 + comorb_malnutrition + 
                comorb_other_GI_notLiver_K_excludesK70K77 +
                comorb_chronic_kidney_disease_N18 +
                comorb_nicotine_dependence_F17 + 
                comorb_heart_failure_I50, data = dtrain,
              Hess = TRUE)


glm_probs = predict(olr_fit, test)

table(test$COVIDseverity, glm_probs)
mean(as.character(olr_df[-train,]$COVIDseverity) != as.character(COVIDseverity))

```

```{r}
payer_olr <- polr(COVIDseverity ~ payer, data = dtrain, Hess = TRUE)

payer_probs = predict(payer_olr, test)

table(test$COVIDseverity, payer_probs)
mean(as.character(olr_df[-train,]$COVIDseverity) != as.character(COVIDseverity))


```

```{r}

first_enc_labs <- d %>%
  mutate(servicedate = as.Date(d[,"servicedate"]),
         personid = as.factor(personid),
         age_group = as.factor(case_when(age_at_encounter <= 11 ~ "0 - 11",
                               age_at_encounter > 11 & 
                                 age_at_encounter <= 17 ~ "12 - 17",
                               age_at_encounter > 17 &
                                 age_at_encounter < 21 ~ "18 - 20",
                               age_at_encounter > 20 ~ "21 - 25")),
         payer = as.factor(payer)) %>% 
  arrange_at("servicedate") %>%
  distinct(personid, .keep_all = TRUE) %>% 
  mutate(resp_rate_lev = as.factor(case_when((age_at_encounter <= 0 & 
                                      respiratoryrate < 30)|
                                    ((age_at_encounter == 1 | 
                                        age_at_encounter== 2) & 
                                       respiratoryrate < 24)|
                                    (age_at_encounter >= 3 & 
                                       age_at_encounter <= 5 & 
                                       respiratoryrate < 22)|
                                    (age_at_encounter >= 6 &
                                       age_at_encounter <= 12 &
                                       respiratoryrate < 18)|
                                    (age_at_encounter >= 13 &
                                       respiratoryrate < 12) ~ "Low",
                                    (age_at_encounter <= 0 & 
                                       respiratoryrate >= 30 &
                                       respiratoryrate <= 60)|
                                    ((age_at_encounter == 1 | 
                                        age_at_encounter == 2) &
                                       respiratoryrate >= 24 &
                                       respiratoryrate <= 40)|
                                    (age_at_encounter >= 3 &
                                       age_at_encounter <= 5 &
                                       respiratoryrate >= 22 &
                                       respiratoryrate <= 34)|
                                    (age_at_encounter >= 6 &
                                       age_at_encounter <= 12 &
                                       respiratoryrate >= 18 &
                                       respiratoryrate <= 30)|
                                    (age_at_encounter >= 13 &
                                       respiratoryrate >= 12 &
                                       respiratoryrate <= 16) ~ "Normal",
                                    (age_at_encounter <= 0 &
                                       respiratoryrate > 60)|
                                    ((age_at_encounter == 1 |
                                        age_at_encounter == 2) &
                                       respiratoryrate > 40)|
                                    (age_at_encounter >= 3 &
                                       age_at_encounter <= 5 &
                                       respiratoryrate > 34)|
                                    (age_at_encounter >= 6 &
                                       age_at_encounter <= 12 &
                                       respiratoryrate > 30)|
                                    (age_at_encounter >= 13 &
                                       respiratoryrate > 16) ~ "High")),
  ord_spo2 = as.factor(case_when(spo2 >= 95 ~ 0,
                       spo2 < 95 ~ 1,
                       is.na(spo2) ~ 0)),
  ord_lymp = as.factor(case_when(lymphocytes_1000_per_uL >= 1 &
                       lymphocytes_1000_per_uL <= 4.8 ~ 0,
                       lymphocytes_1000_per_uL < 1 |                                                     lymphocytes_1000_per_uL > 4.8 ~ 1,
                       is.na(lymphocytes_1000_per_uL) ~ 0)), 
  ord_ALP = as.factor(case_when(ALP_U_per_L >= 44 &
                      ALP_U_per_L <= 147 ~ 0,
                      ALP_U_per_L < 44 | ALP_U_per_L > 147 ~ 1,
                      is.na(ALP_U_per_L) ~ 0)), 
  ord_ALT = as.factor(case_when(ALT_U_per_L >= 10 & ALT_U_per_L <= 40 ~ 0,
                      ALT_U_per_L < 10 | ALT_U_per_L > 40 ~ 1,
                      is.na(ALT_U_per_L) ~ 0)),
  ord_LDH = as.factor(case_when((age_at_encounter < 18 & LDH_U_per_L >= 160 &
                         LDH_U_per_L <= 450)|(age_at_encounter >= 18 &
                                                LDH_U_per_L >= 140 &
                                                LDH_U_per_L <= 280) ~ 0, 
                         (age_at_encounter  < 18 & (LDH_U_per_L <  260 | 
                                                      LDH_U_per_L >  450)) | 
                           (age_at_encounter  >= 18 & (LDH_U_per_L <  140 | 
                                                         LDH_U_per_L >  280)) ~
                           1, 
                         is.na(LDH_U_per_L) ~ 0))) %>% 
  left_join(total_comorb_df, by = "encounterid")
```


```{r}
train_2 <-  sample(nrow(first_enc_labs), size = floor(.80 * 
                                                        nrow(first_enc_labs)))

train_final <- first_enc_labs[train_2,]
test_final <- first_enc_labs[-train_2,]

final_fit <- polr(COVIDseverity ~ comorb_bronchiectasis_J47 +
                comorb_resp_failure_J96 + comorb_malnutrition  + ord_spo2 + 
                  ord_lymp + ord_LDH + resp_rate_lev + ord_ALP + ord_ALT + 
                  comorb_chronic_kidney_disease_N18 + 
                  comorb_nicotine_dependence_F17 +
                  comorb_other_GI_notLiver_K_excludesK70K77 + age_group + payer 
                + comorb_heart_failure_I50 + comorb_asthma_J45 +
                  comorb_copd_J40_J44 +
                  comorb_other_nutritional_deficiencies_E50_E64, 
                data = train_final, Hess = TRUE)


final_probs = predict(final_fit, test_final)

table(test_final$COVIDseverity, final_probs)
mean(as.character(test_final$COVIDseverity) != as.character(COVIDseverity))

```


```{r}
fel_bin <- d %>%
  mutate(COVIDseverity = as.factor(case_when(COVIDseverity > 0 ~ 1,
                                             COVIDseverity == 0 ~ 0)),
         servicedate = as.Date(d[,"servicedate"]),
         personid = as.factor(personid),
         age_group = as.factor(case_when(age_at_encounter <= 11 ~ "0 - 11",
                               age_at_encounter > 11 & 
                                 age_at_encounter <= 17 ~ "12 - 17",
                               age_at_encounter > 17 &
                                 age_at_encounter < 21 ~ "18 - 20",
                               age_at_encounter > 20 ~ "21 - 25")),
         payer = as.factor(payer)) %>% 
  arrange_at("servicedate") %>%
  distinct(personid, .keep_all = TRUE) %>% 
  mutate(resp_rate_lev = as.factor(case_when((age_at_encounter <= 0 & 
                                      respiratoryrate < 30)|
                                    ((age_at_encounter == 1 | 
                                        age_at_encounter== 2) & 
                                       respiratoryrate < 24)|
                                    (age_at_encounter >= 3 & 
                                       age_at_encounter <= 5 & 
                                       respiratoryrate < 22)|
                                    (age_at_encounter >= 6 &
                                       age_at_encounter <= 12 &
                                       respiratoryrate < 18)|
                                    (age_at_encounter >= 13 &
                                       respiratoryrate < 12) ~ "Low",
                                    (age_at_encounter <= 0 & 
                                       respiratoryrate >= 30 &
                                       respiratoryrate <= 60)|
                                    ((age_at_encounter == 1 | 
                                        age_at_encounter == 2) &
                                       respiratoryrate >= 24 &
                                       respiratoryrate <= 40)|
                                    (age_at_encounter >= 3 &
                                       age_at_encounter <= 5 &
                                       respiratoryrate >= 22 &
                                       respiratoryrate <= 34)|
                                    (age_at_encounter >= 6 &
                                       age_at_encounter <= 12 &
                                       respiratoryrate >= 18 &
                                       respiratoryrate <= 30)|
                                    (age_at_encounter >= 13 &
                                       respiratoryrate >= 12 &
                                       respiratoryrate <= 16) ~ "Normal",
                                    (age_at_encounter <= 0 &
                                       respiratoryrate > 60)|
                                    ((age_at_encounter == 1 |
                                        age_at_encounter == 2) &
                                       respiratoryrate > 40)|
                                    (age_at_encounter >= 3 &
                                       age_at_encounter <= 5 &
                                       respiratoryrate > 34)|
                                    (age_at_encounter >= 6 &
                                       age_at_encounter <= 12 &
                                       respiratoryrate > 30)|
                                    (age_at_encounter >= 13 &
                                       respiratoryrate > 16) ~ "High")),
  ord_spo2 = as.factor(case_when(spo2 >= 95 ~ 0,
                       spo2 < 95 ~ 1,
                       is.na(spo2) ~ 0)),
  ord_lymp = as.factor(case_when(lymphocytes_1000_per_uL >= 1 &
                       lymphocytes_1000_per_uL <= 4.8 ~ 0,
                       lymphocytes_1000_per_uL < 1 |                                                     lymphocytes_1000_per_uL > 4.8 ~ 1,
                       is.na(lymphocytes_1000_per_uL) ~ 0)), 
  ord_ALP = as.factor(case_when(ALP_U_per_L >= 44 &
                      ALP_U_per_L <= 147 ~ 0,
                      ALP_U_per_L < 44 | ALP_U_per_L > 147 ~ 1,
                      is.na(ALP_U_per_L) ~ 0)), 
  ord_ALT = as.factor(case_when(ALT_U_per_L >= 10 & ALT_U_per_L <= 40 ~ 0,
                      ALT_U_per_L < 10 | ALT_U_per_L > 40 ~ 1,
                      is.na(ALT_U_per_L) ~ 0)),
  ord_LDH = as.factor(case_when((age_at_encounter < 18 & LDH_U_per_L >= 160 &
                         LDH_U_per_L <= 450)|(age_at_encounter >= 18 &
                                                LDH_U_per_L >= 140 &
                                                LDH_U_per_L <= 280) ~ 0, 
                         (age_at_encounter  < 18 & (LDH_U_per_L <  260 | 
                                                      LDH_U_per_L >  450)) | 
                           (age_at_encounter  >= 18 & (LDH_U_per_L <  140 | 
                                                         LDH_U_per_L >  280)) ~
                           1, 
                         is.na(LDH_U_per_L) ~ 0)))
```


```{r}
binlr <- glm(COVIDseverity ~ comorb_bronchiectasis_J47 +
                  comorb_resp_failure_J96 + comorb_malnutrition  + ord_spo2 + 
                  ord_lymp + ord_LDH + resp_rate_lev + ord_ALP + ord_ALT + 
                  comorb_chronic_kidney_disease_N18 + 
                  comorb_nicotine_dependence_F17 +
                  comorb_other_GI_notLiver_K_excludesK70K77 + age_group + 
                  payer + comorb_heart_failure_I50 + comorb_asthma_J45 +
                  comorb_copd_J40_J44 +
                  comorb_other_nutritional_deficiencies_E50_E64, 
    data = train_final, family = "binomial")

predicted <- predict(binlr, test_final, type = "response")

test_final$default <- ifelse(test_final$default == "Yes", 1, 0)
```

```{r}
total_comorb_df <- dplyr::select(d, contains("comorb")) %>% 
  mutate_if(is.factor, as.numeric) %>% 
  - 1
  
total_comorb_df <- total_comorb_df %>% 
  mutate(total_comorb_by_patient = c(rowSums(total_comorb_df)), 
         encounterid = d[,"encounterid"]) %>% 
  dplyr::select("encounterid", "total_comorb_by_patient")

fel_bin <- d %>%
  mutate(COVIDseverity = as.factor(case_when(COVIDseverity > 0 ~ 1,
                                             COVIDseverity == 0 ~ 0)),
         servicedate = as.Date(d[,"servicedate"]),
         personid = as.factor(personid),
         age_group = as.factor(case_when(age_at_encounter <= 11 ~ "0 - 11",
                               age_at_encounter > 11 & 
                                 age_at_encounter <= 17 ~ "12 - 17",
                               age_at_encounter > 17 &
                                 age_at_encounter < 21 ~ "18 - 20",
                               age_at_encounter > 20 ~ "21 - 25")),
         payer = as.factor(payer)) %>% 
  arrange_at("servicedate") %>%
  distinct(personid, .keep_all = TRUE) %>% 
  mutate(resp_rate_lev = as.factor(case_when((age_at_encounter <= 0 & 
                                      respiratoryrate < 30)|
                                    ((age_at_encounter == 1 | 
                                        age_at_encounter== 2) & 
                                       respiratoryrate < 24)|
                                    (age_at_encounter >= 3 & 
                                       age_at_encounter <= 5 & 
                                       respiratoryrate < 22)|
                                    (age_at_encounter >= 6 &
                                       age_at_encounter <= 12 &
                                       respiratoryrate < 18)|
                                    (age_at_encounter >= 13 &
                                       respiratoryrate < 12) ~ "Low",
                                    (age_at_encounter <= 0 & 
                                       respiratoryrate >= 30 &
                                       respiratoryrate <= 60)|
                                    ((age_at_encounter == 1 | 
                                        age_at_encounter == 2) &
                                       respiratoryrate >= 24 &
                                       respiratoryrate <= 40)|
                                    (age_at_encounter >= 3 &
                                       age_at_encounter <= 5 &
                                       respiratoryrate >= 22 &
                                       respiratoryrate <= 34)|
                                    (age_at_encounter >= 6 &
                                       age_at_encounter <= 12 &
                                       respiratoryrate >= 18 &
                                       respiratoryrate <= 30)|
                                    (age_at_encounter >= 13 &
                                       respiratoryrate >= 12 &
                                       respiratoryrate <= 16) ~ "Normal",
                                    (age_at_encounter <= 0 &
                                       respiratoryrate > 60)|
                                    ((age_at_encounter == 1 |
                                        age_at_encounter == 2) &
                                       respiratoryrate > 40)|
                                    (age_at_encounter >= 3 &
                                       age_at_encounter <= 5 &
                                       respiratoryrate > 34)|
                                    (age_at_encounter >= 6 &
                                       age_at_encounter <= 12 &
                                       respiratoryrate > 30)|
                                    (age_at_encounter >= 13 &
                                       respiratoryrate > 16) ~ "High")),
  ord_spo2 = as.factor(case_when(spo2 >= 95 ~ 0,
                       spo2 < 95 ~ 1,
                       is.na(spo2) ~ 0)),
  ord_lymp = as.factor(case_when(lymphocytes_1000_per_uL >= 1 &
                       lymphocytes_1000_per_uL <= 4.8 ~ 0,
                       lymphocytes_1000_per_uL < 1 |                                                     lymphocytes_1000_per_uL > 4.8 ~ 1,
                       is.na(lymphocytes_1000_per_uL) ~ 0)), 
  ord_ALP = as.factor(case_when(ALP_U_per_L >= 44 &
                      ALP_U_per_L <= 147 ~ 0,
                      ALP_U_per_L < 44 | ALP_U_per_L > 147 ~ 1,
                      is.na(ALP_U_per_L) ~ 0)), 
  ord_ALT = as.factor(case_when(ALT_U_per_L >= 10 & ALT_U_per_L <= 40 ~ 0,
                      ALT_U_per_L < 10 | ALT_U_per_L > 40 ~ 1,
                      is.na(ALT_U_per_L) ~ 0)),
  ord_LDH = as.factor(case_when((age_at_encounter < 18 & LDH_U_per_L >= 160 &
                         LDH_U_per_L <= 450)|(age_at_encounter >= 18 &
                                                LDH_U_per_L >= 140 &
                                                LDH_U_per_L <= 280) ~ 0, 
                         (age_at_encounter  < 18 & (LDH_U_per_L <  260 | 
                                                      LDH_U_per_L >  450)) | 
                           (age_at_encounter  >= 18 & (LDH_U_per_L <  140 | 
                                                         LDH_U_per_L >  280)) ~
                           1, 
                         is.na(LDH_U_per_L) ~ 0))) %>% 
  left_join(total_comorb_df, by = "encounterid")

```


```{r}
glm(COVIDseverity ~ total_comorb_by_patient + age_group, data = first_enc,
    family = "binomial") %>% 
  summary()

table(first_enc$total_comorb_by_patient)/length(first_enc$total_comorb_by_patient)
```
Having one comorbidity will increase the log odds of needing oxygen therapy by
0.033. Having two comorbidities will increase the log odds of needing oxygen
therapy by 0.353. Having three or more comorbidities increases the log odds
of needing oxygen therapy by 1.415. 


