---
title: "Week 3 - CF Data"
author: "Terrell D Lemons"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This Graph is a representation of COMORBITIES of a patient with Cystic Fibrosis and the percentage of those patients were affected by those COMORBITIES. 


```{r "CF Dataset 1", echo=FALSE}
library(Hmisc)
library(tidyverse)
library(viridis)

tibble("cfCOVIDgroup6.RDATA")

names(d)

demo_3 <- tribble(
  ~COMORBIDITY,         ~PATIENTS,
  "Type I Diabete",       216/9337,
  "Type II Diabete",       310/9337,

  "Bi-Pap Cpap Non-Invasive",       66/9337,
  "Invasive Mech Ventilatio",       260/9337,

  "Malnutrition",       440/9337,
  "Other Nutritional Deficiencies",      235/9337,
  
  "Chronic Kidney Disease",       42/9337,
  "Liver Disease",       265/9337,
  "Obesity Overweight",      694/9337,
  "Other GI Not Liver",      3251/9337,
)

tdl_3 <- demo_3 %>%
  ggplot(aes(x = COMORBIDITY, y = PATIENTS, fill=COMORBIDITY, position = "dodge")) +
  labs(title="PRECENTAGES NOT AFFECTED BY COMORBIDITIES") +
  guides(fill=guide_legend("Non-Respiratory")) +
  stat_summary(fun = mean, geom = "bar") + 
  
  geom_bar(stat="identity") +
  scale_fill_manual(values=c("#D90EF0",
                             "#C30DD8",
                             "#AD0CC0",
                             "#980AA8",
                             "#820990",
                             "#6C0778",
                             "#570660",
                             "#410448",
                             "#2B0330",
                             "#160118")) +
  
  stat_summary(fun.data=mean_cl_normal, geom="errorbar", width = 0.25) + 
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(vjust = +3)) +

scale_y_continuous(labels = scales::percent)

tdl_3
```


## The Variables: Comorbidities used to gauge the effects of two or more underlining factors contributed to COVID-19


```{r}
install.packages("viridis")  # Install
library("viridis")           # Load

library("Hmisc")
library("tidyverse")

tibble("cfCOVIDgroup6.RDATA")

demo_3 <- tribble(
  ~COMORBIDITY,         ~PATIENTS,
  "Heart Disease",       6/9337,
  "Heart Failure",       25/9337,

  "Type I Diabete",       216/9337,
  "Type II Diabete",       310/9337,

  "Bi-Pap Cpap Non-Invasive",       66/9337,
  "Invasive Mech Ventilatio",       260/9337,

  "Malnutrition",       440/9337,
  "Other Nutritional Deficiencies",      235/9337,
)

ggplot(demo_3, aes(x = COMORBIDITY, y = PATIENTS, fill = COMORBIDITY)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, hjust = 3))
```


## The Variables: Comorbidities used to gauge the effects of two or more underlining factors contributed to COVID-19


```{r "CF Dataset 2", echo=FALSE}
library("Hmisc")
library("tidyverse")
library("ggplot2")

tibble("cfCOVIDgroup6.RDATA")

demo_2 <- tribble(
  ~COMORBIDITY,         ~PATIENTS,
  "Chronic Kidney Disease",       42/9337,
  "Liver Disease",       265/9337,
  "Obesity Overweight",      694/9337,
  "Other GI Not Liver",      3251/9337,
)

ggplot(demo_2, aes(x = COMORBIDITY, y = PATIENTS, fill = PATIENTS)) +
  geom_bar(stat = "identity") +
  geom_col(position = position_dodge(0.1), colour = "black", size = 0.5) +
  coord_cartesian(ylim = c(0, NA)) +
  theme_bw()
```

## The Variables: Comorbidities used to gauge the effects of two or more underlining factors contributed to COVID-19

```{r "CF Dataset 3", echo=FALSE}
library("Hmisc")
library("tidyverse")

tibble("cfCOVIDgroup6.RDATA")

demo_1 <- tribble(
  ~COMORBIDITY,         ~PATIENTS,
  "Asthma",      2162/9337,
  "Bronchiectasis",  308/9337,
  "COPD",      678/9337,
  "Lung Transplant",       23/9337,
  "Nasal Polyps",      76/9337,
  "Pneumothorax",    27/9337,
  "Resp Failure",       369/9337,
)

ggplot(demo_1, aes(COMORBIDITY, PATIENTS, fill=rownames(demo_1))) +
  geom_bar(stat = "identity") + guides(fill=FALSE) +
  # geom_col(colour = "black") + 
  # geom_point(colour = "black") +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.25) +  # nolint
  theme(legend.position = "bottom")
```


```{r}
library("Hmisc")
library("tidyverse")
library("gridExtra")
library("ggpubr")
library("ggridges")

tibble("cfCOVIDgroup6.RDATA")

demo_1 <- tribble(
  ~COMORBIDITY,         ~PATIENTS,
  "Asthma",      2162/9337,
  "Bronchiectasis",  308/9337,
  "COPD",      678/9337,
  "Lung Transplant",       23/9337,
  "Nasal Polyps",      76/9337,
  "Pneumothorax",    27/9337,
  "Resp Failure",       369/9337,
)

demo_2 <- tribble(
  
  # COMORBITIES AND PATIENTS
  ~COMORBIDITY,         ~PATIENTS,
  
  "Heart Disease",       6/9337,
  "Heart Failure",       25/9337,

  "Type I Diabete",       216/9337,
  "Type II Diabete",       310/9337,

  "Chronic Kidney Disease",       42/9337,
  "Liver Disease",       265/9337,
  "Obesity Overweight",      694/9337,
  "Other GI Not Liver",      3251/9337,
)

demo_9 <- tribble(
  
  # COMORBITIES AND PATIENTS
  ~COMORBIDITY,         ~PATIENTS,
  
  "Heart Disease",       6/9337,
  "Heart Failure",       25/9337,

  "Type I Diabete",       216/9337,
  "Type II Diabete",       310/9337,

  "Chronic Kidney Disease",       42/9337,
  "Liver Disease",       265/9337,
  "Obesity Overweight",      694/9337,
  "Other GI Not Liver",      3251/9337,
)

tdl_1 <- demo_1 %>%
  ggplot(aes(COMORBIDITY, PATIENTS, fill=rownames(demo_1))) +
  labs(title="COMORBIDITY (%) BY PATIENTS") +
    geom_bar(stat = "identity") + 
    # LEGEND  
    guides(fill="none") +
    scale_fill_manual(values=c("#BD1359",
                               "#AA1150",
                               "#970F47",
                               "#840D3E",
                               "#710B35",
                               "#5E092C",
                               "#4B0824",
                               "#39061B",
                               "#260412",
                               "#130209")) +
  
    geom_col(colour = "black") + 
    # geom_point(colour = "black") +
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.25) + 
    theme(axis.text.x = element_text(vjust = -2.75),
          axis.title.x = element_blank()) +
          #axis.text.x=element_blank()) +
          # axis.ticks.x=element_blank()) +
    # guides(x = guide_axis(n.dodge = 2)) +
    scale_y_continuous(labels = label_percent(accuracy = 1))

tdl_2 <- demo_2 %>%
  ggplot(aes(x = COMORBIDITY, y = PATIENTS, fill = COMORBIDITY)) +
    geom_bar(stat = "identity") +
    # scale_fill_manual(values=c("#6E13BD",
                               #"#6311AA",
                               #"#580F97",
                               #"#4D0D84",
                               #"#420B71",
                               #"#37095E",
                               #"#2C084B",
                               #"#210639",
                               #"#160426",
                               #"#0B0213")) +
  
    geom_col(position = position_dodge(0.1), 
             colour = "black", 
             size = 0.5) +
  
    coord_cartesian(ylim = c(0, NA)) +
  
    theme_bw() +
    theme(
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.y = element_text(vjust = +3),
          axis.title.x = element_blank()) +
    guides(x = guide_axis(n.dodge = 2)) +
    scale_y_continuous(labels = label_percent(accuracy = 1))

# grid.arrange(tdl_1, arrangeGrob(tdl_2, tdl_2, ncol=2), nrow = 3)
  




tdl_9 <- demo_9 %>%
  ggplot(aes(iris, x = PATIENTS, y = COMORBIDITY, fill = PATIENTS)) +
  geom_density_ridges(aes(fill = COMORBIDITY)) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800", "#FC4E07", "#00AFBB", "#E7B800", "#FC4E07", "#00AFBB", "#E7B800", "#FC4E07"))


grid.arrange(tdl_1, arrangeGrob(tdl_2, tdl_9, nrow=2), nrow = 1)
      
# grid.arrange(tdl_1, tdl_2, nrow = 2, ncol = 2)
```



```{r}
#remove.packages("ggplot2") # Unisntall ggplot
#install.packages("ggplot2") # Install it again
library(ggplot2)
library(Hmisc)
library(tidyverse)
library(gridExtra)

tibble("cfCOVIDgroup6.RDATA")

x <- 1:10                                     # Create example 

y1 <- c(3, 1, 5, 2, 3, 8, 4, 7, 6, 9)         # Create data

y2 <- c(5, 1, 4, 6, 2, 3, 7, 8, 2, 8)         # Create more data 
y3 <- c(3, 3, 3, 3, 4, 4, 5, 5, 7, 7)

plot(x, y1, type = "b",                       # Change title & axis   
  # labels
     main = "SEVERITY WITH COVID",
     xlab = "1 to 10 Severity",
     ylab = "Frequency",
     col = "blue",
     pch = 16,
     # lwd = 5,                               # Thickness of line
)

# Add second & third lines
lines(x, y2, type = "b", col = "red", pch = 15)         
lines(x, y3, type = "b", col = "green", pch = 8)  

legend(
  "topleft",                             # Add legend to plot
    legend = c("Men", "Women", "Children"),
    col = c("blue", "red", "green"),
    pch = c(16, 15, 8),
    lty = 1
)

```


```{r}
library(ggplot2)
library(Hmisc)
library(tidyverse)
library(gridExtra)
library(scales)

tibble("cfCOVIDgroup6.RDATA")

names(d)

#select(d, gender, age_at_encounter, zip_code, tenant)

#d <- data.frame(zip_code = c("01234", "39999", "59999", "79999", "99999"),             age_at_encounter = c(3, 7, 11, 15, 19))

ggplot(d, aes(zip_code, age_at_encounter) ) +
  labs(title="COVID ENCOUNTERS BY ZIP CODE AND AGE",
    x = "Zip Code", 
    y = "Age of Encounter") +
  
  #geom_point() +
  # xlim(00, 09) + 
  # ylim(18,20) +
  
  geom_smooth(method = "lm", se = FALSE, aes(color = "green")) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = FALSE, aes(color = "red")) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, aes(color = "blue")) +
  # scale_color_manual(values = c("green", "red", "blue") ) +

  scale_color_identity(
    name = "Patients:",
    breaks = c("green", "red", "blue"),
    labels = c("Children", "Female", "Male"),
    guide = "legend") +
  theme(legend.position = "bottom",
        axis.title.y = element_text(vjust = +3),
        axis.title.x = element_text(vjust = -0.75)) + 
  scale_x_continuous(labels = label_number(accuracy = 1))

```

