---
title: "Week 3 - CF Data"
author: "Terrell D Lemons"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This Graph is a representation of COMORBITIES of a patient with Cystic Fibrosis and the percentage of those patients were affected by those COMORBITIES. 

```{r "CF Dataset 1", echo=FALSE}
library(Hmisc)
library(tidyverse)
library(viridis)

tibble("cfCOVIDgroup6.RDATA")

names(d)

demo_3 <- tribble(
  ~COMORBIDITY,                          ~PATIENTS,
  "Type I Diabete",                      216/9337,
  "Type II Diabete",                     310/9337,

  "Bi-Pap Cpap Non-Invasive",            66/9337,
  "Invasive Mech Ventilatio",            260/9337,

  "Malnutrition",                        440/9337,
  "Other Nutritional Deficiencies",      235/9337,
  
  "Chronic Kidney Disease",              42/9337,
  "Liver Disease",                       265/9337,
  "Obesity Overweight",                  694/9337,
  "Other GI Not Liver",                  3251/9337,
)

tdl_3 <- demo_3 %>%
  ggplot(aes(x = COMORBIDITY, y = PATIENTS, fill=COMORBIDITY, position = "dodge")) +
  labs(title="PRECENTAGES NOT AFFECTED BY COMORBIDITIES") +
  
  guides(fill=guide_legend("Non-Respiratory")) +
  stat_summary(fun = mean, geom = "bar") + 
  
  geom_bar(stat="identity") +
  scale_fill_manual(values=c("#D90EF0",
                             "#C30DD8",
                             "#AD0CC0",
                             "#980AA8",
                             "#820990",
                             "#6C0778",
                             "#570660",
                             "#410448",
                             "#2B0330",
                             "#160118")) +
  
  stat_summary(fun.data=mean_cl_normal, geom="errorbar", width = 0.25) + 
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(vjust = +3)) +

  scale_y_continuous(labels = scales::percent)

tdl_3
```


## The Variables: Comorbidities used to gauge the effects of two or more underlining factors contributed to COVID-19


```{r}
library("viridis")           # Load
library("Hmisc")
library("tidyverse")

tibble("cfCOVIDgroup6.RDATA")

demo_3 <- tribble(
  ~COMORBIDITY,                          ~PATIENTS,
  "Heart Disease",                       6/9337,
  "Heart Failure",                       25/9337,

  "Type I Diabete",                      216/9337,
  "Type II Diabete",                     310/9337,

  "Bi-Pap Cpap Non-Invasive",            66/9337,
  "Invasive Mech Ventilatio",            260/9337,

  "Malnutrition",                        440/9337,
  "Other Nutritional Deficiencies",      235/9337,
)

ggplot(demo_3, aes(x = COMORBIDITY, y = PATIENTS, fill = COMORBIDITY)) +
  labs(title = "OTHER COMORBIDITIES") +
  
  geom_bar(stat = "identity") +
  theme(
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(vjust = +3)) +
  
  scale_y_continuous(labels = scales::percent)
```


## The Variables: Comorbidities used to gauge the effects of two or more underlining factors contributed to COVID-19


```{r "CF Dataset 2", echo=FALSE}
library("Hmisc")
library("tidyverse")
library("ggplot2")

tibble("cfCOVIDgroup6.RDATA")

demo_2 <- tribble(
  ~COMORBIDITY,                   ~PATIENTS,
  "Chronic Kidney Disease",       42/9337,
  "Liver Disease",                265/9337,
  "Obesity Overweight",           694/9337,
  "Other GI Not Liver",           3251/9337,
)

ggplot(demo_2, aes(x = COMORBIDITY, y = PATIENTS, fill = COMORBIDITY))  +
  labs(title = "GI COMORBIDITIES") +
  
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_text(vjust = -1.75),
        axis.title.y = element_text(vjust = +3)) +
  
  scale_y_continuous(labels = scales::percent)
```

## The Variables: Comorbidities used to gauge the effects of two or more underlining factors contributed to COVID-19

```{r "CF Dataset 3", echo=FALSE}
library("Hmisc")
library("tidyverse")

tibble("cfCOVIDgroup6.RDATA")

demo_1 <- tribble(
  ~COMORBIDITY,         ~PATIENTS,
  "Asthma",             2162/9337,
  "Bronchiectasis",     308/9337,
  
  "COPD",               678/9337,
  "Lung Transplant",    23/9337,
  
  "Nasal Polyps",       76/9337,
  "Pneumothorax",       27/9337,
  
  "Resp Failure",       369/9337,
  "Inv Mech Ven",       260/9337,
)

ggplot(demo_1, aes(COMORBIDITY, PATIENTS, fill=COMORBIDITY,)) +
  
  ## MAIN TITLE, TITLE, LEGEND
  labs(title="RESPIRATORY COMORBIDITIES") +
    #x = "Zip Code", 
    #y = "Age of Encounter") +
  
  guides(fill = guide_legend("")) +
  
  geom_bar(stat = "identity") + 

  stat_summary(
    fun.data = mean_cl_normal, 
    geom = "errorbar", 
    width = 0.25) +  # nolint

  theme(
      legend.position = "bottom",
      axis.title.x = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.y = element_text(vjust = +3)) +
  
  scale_y_continuous(labels = scales::percent)
```


```{r}
library("Hmisc")
library("tidyverse")
library("gridExtra")
library("ggpubr")
library("ggridges")

tibble("cfCOVIDgroup6.RDATA")

demo_1 <- tribble(
  ~COMORBIDITY,         ~PATIENTS,
  "Asthma",             2162/9337,
  "Bronchiectasis",     308/9337,
  "COPD",               678/9337,
  "Lung Transplant",    23/9337,
  "Nasal Polyps",       76/9337,
  "Pneumothorax",       27/9337,
  "Resp Failure",       369/9337,
)

demo_2 <- tribble(
  
  # COMORBITIES AND PATIENTS
  ~COMORBIDITY,                   ~PATIENTS,
  
  "Heart Disease",                6/9337,
  "Heart Failure",                25/9337,

  "Type I Diabete",               216/9337,
  "Type II Diabete",              310/9337,

  "Chronic Kidney Disease",       42/9337,
  "Liver Disease",                265/9337,
  "Obesity Overweight",           694/9337,
  "Other GI Not Liver",           3251/9337,
)

demo_9 <- tribble(
  
  # COMORBITIES AND PATIENTS
  ~COMORBIDITY,                   ~PATIENTS,
  
  "Heart Disease",                6/9337,
  "Heart Failure",                25/9337,

  "Type I Diabete",               216/9337,
  "Type II Diabete",              310/9337,

  "Chronic Kidney Disease",       42/9337,
  "Liver Disease",                265/9337,
  "Obesity Overweight",           694/9337,
  "Other GI Not Liver",           3251/9337,
)

tdl_1 <- demo_1 %>%
  ggplot(aes(COMORBIDITY, PATIENTS, fill = COMORBIDITY)) +
  labs(title="COMORBIDITY (%) BY PATIENTS") +
    geom_bar(stat = "identity") + 
    
    # LEGEND  
    guides(fill = guide_legend("")) +
  
    geom_col(colour = "black") + 
  
    stat_summary(fun.data = mean_cl_normal, 
                 geom = "errorbar", 
                 width = 0.25) + 
  
    theme(
        legend.position = "bottom",
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(vjust = +3),
        axis.title.x = element_blank()) +
      
    scale_y_continuous(labels = scales::percent)


tdl_2 <- demo_2 %>%
  ggplot(aes(x = COMORBIDITY, y = PATIENTS, fill = COMORBIDITY)) +
    geom_bar(stat = "identity") +
    scale_fill_manual(values=c("6E13BD",
                               "#6311AA",
                               "#580F97",
                               "#4D0D84",
                               "#420B71",
                               "#37095E",
                               "#2C084B",
                               "#210639",
                               "#160426",
                               "#0B0213")
                              ) +
  
    coord_cartesian(ylim = c(0, NA)) +
  
    theme_bw() +
    theme(
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.y = element_text(vjust = +3),
          axis.title.x = element_blank()) +
  
    guides(x = guide_axis(n.dodge = 2)) +
   
    scale_y_continuous(labels = scales::percent)


tdl_9 <- demo_9 %>%
  ggplot(aes(iris, x = PATIENTS, y = COMORBIDITY, fill = PATIENTS)) +
  geom_density_ridges(aes(fill = COMORBIDITY)) +
  scale_fill_manual(values = c("#00AFBB", 
                               "#E7B800", 
                               "#FC4E07", 
                               "#00AFBB", 
                               "#E7B800", 
                               "#FC4E07", 
                               "#00AFBB", 
                               "#E7B800", 
                               "#FC4E07")
                              ) +
  
   theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(vjust = +3),
        axis.title.x = element_blank()) +
  
  scale_y_continuous(labels = scales::percent)
  
grid.arrange(tdl_1, tdl_2, nrow = 2, ncol = 1)
```



```{r}

library(ggplot2)
library(Hmisc)
library(tidyverse)
library(gridExtra)

tibble("cfCOVIDgroup6.RDATA")

x <- 1:10                                     # Create example 

y1 <- c(3, 1, 5, 2, 3, 8, 4, 7, 6, 9)         # Create data

y2 <- c(5, 1, 4, 6, 2, 3, 7, 8, 2, 8)         # Create more data 
y3 <- c(3, 3, 3, 3, 4, 4, 5, 5, 7, 7)

plot(x, y1, type = "b",                       # Change title & axis   
  # labels
     main = "SEVERITY WITH COVID",
     xlab = "1 to 10 Severity",
     ylab = "Frequency",
     col = "blue",
     pch = 16,
     # lwd = 5,                               # Thickness of line
)

# Add second & third lines
lines(x, y2, type = "b", col = "red", pch = 15)         
lines(x, y3, type = "b", col = "green", pch = 8)  

legend(
  "topleft",                                  # Add legend to plot
    legend = c("Men", "Women", "Children"),
    col = c("blue", "red", "green"),
    pch = c(16, 15, 8),
    lty = 1
)

```


```{r}
library(ggplot2)
library(Hmisc)
library(tidyverse)
library(gridExtra)
library(scales)

tibble("cfCOVIDgroup6.RDATA")

names(d)

ggplot(d, aes(zip_code, age_at_encounter) ) +
  labs(title="COVID ENCOUNTERS BY ZIP CODE AND AGE",
    x = "Zip Code", 
    y = "Age of Encounter") +
  
  geom_smooth(method = "lm", se = FALSE, aes(color = "green")) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = FALSE, aes(color = "red")) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, aes(color = "blue")) +

  scale_color_identity(
    name = "Patients:",
    breaks = c("green", "red", "blue"),
    labels = c("Children", "Female", "Male"),
    guide = "legend") +
  
  theme(legend.position = "bottom",
        axis.title.y = element_text(vjust = +3),
        axis.title.x = element_text(vjust = -0.75)) + 
  
  scale_x_continuous(labels = label_number(accuracy = 1))

```

