---
title: "06-29 Research Draft"
author: "Hester Nguyen"
date: "6/29/2022"
output: pdf_document
---
<<<<<<< HEAD:06-29 Research Draft.Rmd



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r packages}
library(tidyverse)
```

## R Markdown





```{r}

#library(Microsoft365R)
#rsrch_data <- get_personal_onedrive()
#rsrch_data



glimpse(d)


```

**Hester's EDA**

```{r}
#grouping patients into different kinds of comorbidity that they have
comorb_groups <- d %>% 
  select(personid, age_at_encounter, encounterid, starts_with("comorb"),COVIDseverity) %>% 
  mutate(
         scomorb_J = case_when(
           comorb_copd_J40_J44 == 1|
           comorb_nasal_polyps_J33 == 1|
           comorb_pneumothorax_J93==1|
           comorb_resp_failure_J96 == 1|
           comorb_asthma_J45 == 1 |
           comorb_bronchiectasis_J47 == 1~1,
           comorb_copd_J40_J44 == 0&
           comorb_nasal_polyps_J33 == 0&
           comorb_pneumothorax_J93==0&
           comorb_resp_failure_J96 == 0&
           comorb_asthma_J45 == 0&
           comorb_bronchiectasis_J47 == 0~0
            
         ), scomorb_mal = comorb_malnutrition, 
         scomorb_E = case_when(
           comorb_typeI_diabetes_E10 == 1|
           comorb_obesity_overweight_E66 ==1|
           comorb_typeII_diabetes_E11==1|
           comorb_other_nutritional_deficiencies_E50_E64 == 1~1,
           comorb_typeI_diabetes_E10 == 0&
           comorb_obesity_overweight_E66 ==0&
           comorb_typeII_diabetes_E11==0&
           comorb_other_nutritional_deficiencies_E50_E64 == 0~0
         ),
         scomorb_K = case_when(
           comorb_other_GI_notLiver_K_excludesK70K77 == 1|
           comorb_liver_disease_K70_K77 == 1~1,
           comorb_other_GI_notLiver_K_excludesK70K77 == 0&
           comorb_liver_disease_K70_K77 == 0~0
         ),
         scomorb_I = case_when(
           comorb_hypertensive_heart_disease_I11 == 1|
           comorb_essential_hypertension_I10 == 1|
           comorb_heart_failure_I50==1|
           comorb_ischemic_heart_disease_I20_I25==1~1,
           comorb_hypertensive_heart_disease_I11 == 0&
           comorb_essential_hypertension_I10 == 0&
           comorb_heart_failure_I50==0&
           comorb_ischemic_heart_disease_I20_I25==0~0
         ),
         scomorb_F = case_when(
           comorb_nicotine_dependence_F17 == 1~1,
           comorb_nicotine_dependence_F17== 0~0
           
         ), 
         scomorb_N = case_when(
           comorb_chronic_kidney_disease_N18 == 1~1,
           comorb_chronic_kidney_disease_N18 == 0~0
         ),
         scomorb_Z = case_when(
           comorb_lung_transplant_Z942 == 1~1, 
           comorb_lung_transplant_Z942 == 0~0
         )
         ) %>% 
  select(-starts_with('comorb'))
comorb_groups




```



```{r}
#classifying patient's comorbidities into levels


comorb_type <- 
  mutate(comorb_groups,comorb_types = case_when(
    scomorb_J == 1~'J',
    scomorb_mal ==1~'mal',
    scomorb_E==1~'E',
    scomorb_K ==1 ~'K',
    scomorb_I == 1~'I',
    scomorb_F == 1 ~'F',
    scomorb_N == 1 ~'N',
    scomorb_Z ==1 ~"Z",
    TRUE ~ 'None'
    
  ))
comorb_type
```

```{r}
comorb_type %>% 
  ggplot(aes(x = as.factor(COVIDseverity), , fill = as.factor(comorb_types)))+
  geom_bar(position = "dodge") +
  xlab("COVID Severity") +
  ggtitle("Comorbidity Frequnecy by COVID Severity") +
  scale_fill_discrete(name = "Comorbidity Type", labels = c("Nutritional", 
                                                            "Nicotine Dependence", 
                                                            "Cardio Vascular", 
                                                            "Respiratory",
                                                            "GI", 
                                                            "Nutritional", 
                                                            "Kidney Disease",
                                                            "None"))
```

```{r}

```


