---
title: "Liz Comorb Stuff"
author: "Team 6"
date: '2022-07-01'
output: html_document
---

```{r setup, warning = FALSE, include = FALSE}
<<<<<<< HEAD
#setwd("C:\\Users\\lizvi\\Cal State Fullerton\\Jaynes, Jessica - CSUF CHOC Summer 2022 Research\\Group2")

#load("cfCOVIDgroup6.RDATA")
=======

setwd("")

load("cfCOVIDgroup6.RDATA")
>>>>>>> f0d1d648d25f1c177e15ec0d99962de892904430

library(tidyverse)
library(data.table)
library(gganimate)
<<<<<<< HEAD
#library(gifski)
=======
library(gifski)
library(reshape2)
>>>>>>> f0d1d648d25f1c177e15ec0d99962de892904430

all_comorb <- select(d, contains("comorb"))

glimpse(all_comorb)

attach(d)
```


<<<<<<< HEAD
```{r}
=======
```{r warning = FALSE}
>>>>>>> f0d1d648d25f1c177e15ec0d99962de892904430

all_comorb <- all_comorb %>% mutate_if(is.factor, as.numeric) %>% 
  -1

<<<<<<< HEAD
=======
all_comorb <- all_comorb %>% 
  mutate(total_comorb_by_patient = c(rowSums(all_comorb)), 
         COVIDseverity = d[,73], age_at_encounter = d[,11], service_date = as.Date(d[,9]))

all_comorb %>% 
  summarize(mean_total = mean(total_comorb_by_patient), med_total = median(total_comorb_by_patient), max_total = max(total_comorb_by_patient),
            var_total = var(total_comorb_by_patient))
>>>>>>> f0d1d648d25f1c177e15ec0d99962de892904430
detach(d)

attach(all_comorb)

<<<<<<< HEAD
total_comorb_by_patient <- c(rowSums(all_comorb))

colSums(all_comorb)
=======
total_comorb_by_patient <- c(rowSums(all_comorb[,1:21]))

colSums(all_comorb[,1:21])
>>>>>>> f0d1d648d25f1c177e15ec0d99962de892904430

respiratory_comorb <- all_comorb %>% 
  select(contains("J")) %>% 
  sum()

gi_comorb <- select(all_comorb, contains("K", ignore.case = FALSE)) %>% 
  sum()

nutrition_comorb <- select(all_comorb, contains("E", ignore.case = FALSE), contains("nutrition")) %>% 
  sum()

cardio_comorb <- select(all_comorb, contains("I", ignore.case = FALSE) & 
         !contains("E", ignore.case = FALSE) & 
<<<<<<< HEAD
         !contains("K", ignore.case =  FALSE)) %>% 
=======
         !contains("K", ignore.case =  FALSE) &
         !contains("COVIDseverity")) %>% 
>>>>>>> f0d1d648d25f1c177e15ec0d99962de892904430
  sum()

grouped_comorb <- data.frame(respiratory_comorb, gi_comorb, nutrition_comorb, 
                  cardio_comorb, sum(comorb_chronic_kidney_disease_N18), sum(comorb_hemoptysisR042), sum(comorb_lung_transplant_Z942), 
                  sum(comorb_nicotine_dependence_F17)) %>% 
  colSums() %>% 
  melt(value.name = 'Patients') %>% 
  mutate(comorbidity = c("Respiratory", "GI", "Nutrition", 
                  "Cardiovascular", "Chronic Kidney Disease", "Hemoptysis", "Lung Transplant", "Nicotine Dependence"))



grouped_comorb %>% 
  ggplot(aes(x = comorbidity, y = Patients)) +
  geom_bar(stat = 'identity')

<<<<<<< HEAD


```

```{r}
# ggplot(choc, aes(gdp_per_capita_ppp_constant_2017_international, deaths_cause_all_causes_risk_household_air_pollution_from_solid_fuels_sex_both_age_age_standardized_rate, size = population_historical_estimates, color = entity)) +
#   geom_point(alpha = 0.7, show.legend = FALSE) +
#   scale_x_log10() +
#   facet_wrap(~continent) +
#   labs(title = 'Year: {as.integer(frame_time)}', x = 'GDP per capita', y = 'Age Standardized Rate of Death') +
#   transition_time(year) +
#   ease_aes('linear') +
#   theme_bw() +
#   theme(text = element_text(size = 15))
=======
all_comorb %>% 
  ggplot(aes(x = total_comorb_by_patient)) +
  geom_bar(data = subset(all_comorb, total_comorb_by_patient > 0), fill = "lightsteelblue") +
  scale_x_continuous(breaks = c(1:10)) +
  xlab("Total Number of Comorbidities") +
  ylab("Patients")

all_comorb %>% 
  ggplot(aes(x = total_comorb_by_patient)) +
  geom_bar(fill = "lightsteelblue") +
  scale_x_continuous(breaks = c(0:10)) +
  xlab("Total Number of Comorbidities per Patient") +
  ylab("Patients") +
  facet_wrap(COVIDseverity)


>>>>>>> f0d1d648d25f1c177e15ec0d99962de892904430
```


```{r}
<<<<<<< HEAD
#age and Covid severity

ggplot(d, aes(x = as.factor(age_at_encounter), fill = as.factor(COVIDseverity))) +
  geom_bar(position = position_dodge()) +
  theme_classicf() +
 scale_fill_manual(values=c('lightskyblue1',
                            'sandybrown',
                            'palegreen1',
                            'lightpink'))
```

```{r}
#Race and severity

all_comorb %>% 
  ggplot(aes(x = total_comorb_by_patient)) +
  geom_bar(data = subset(all_comorb, total_comorb_by_patient > 0), fill = "lightsteelblue") +
  scale_x_continuous(breaks = c(1:10)) +
  xlab("Total Number of Comorbidities") +
  ylab("Patients")

all_comorb %>% 
  ggplot(aes(x = total_comorb_by_patient)) +
  geom_bar(fill = "lightsteelblue") +
  scale_x_continuous(breaks = c(0:10)) +
  xlab("Total Number of Comorbidities") +
  ylab("Patients")
```



=======
all_comorb %>% 
  ggplot(aes(x = total_comorb_by_patient, fill = COVIDseverity)) +
  geom_bar() +
  scale_x_continuous(breaks = c(0:10)) +
  xlab("Total Number of Comorbidities per Patient") +
  ylab("Patients") 
 


```


```{r}
all_comorb %>% 
  ggplot(aes(x = COVIDseverity)) +
  geom_bar(fill = "lightsteelblue") +
  xlab("COVID_19 Severity") +
  ylab("Patients") +
  facet_wrap(total_comorb_by_patient)
```


```{r}
ggplot(all_comorb, aes(total_comorb_by_patient, fill = "rainbow")) +
  geom_bar(show.legend = FALSE) +
  scale_x_continuous(breaks = c(0:10)) +
  labs(title = 'Age: {as.integer(frame_time)}', x = 'Total Comorbidities per Patient',
       y = 'Number of Patients') +
  transition_time(age_at_encounter) +
  ease_aes('linear') +
  theme_bw() +
  theme(text = element_text(size = 15))
```


```{r}
ggplot(all_comorb, aes(total_comorb_by_patient, fill = COVIDseverity)) +
  geom_bar(position = "fill") +
  scale_x_continuous(breaks = c(0:10))
```
>>>>>>> f0d1d648d25f1c177e15ec0d99962de892904430

