---
title: "Liz Comorb Stuff"
author: "Team 6"
date: '2022-07-01'
output: html_document
---

```{r setup, warning = FALSE, include = FALSE}

setwd("C:/Users/lizvi/Cal State Fullerton/Jaynes, Jessica - CSUF CHOC Summer 2022 Research/Group2")

load("cfCOVIDgroup6.RDATA")


library(tidyverse)
library(data.table)
library(gganimate)
library(gifski)
library(reshape2)
library(viridis)
library(glmnet)


all_comorb <- dplyr::select(d, contains("comorb"))

glimpse(all_comorb)

attach(d)
```


```{r warning = FALSE}

all_comorb <- all_comorb %>% mutate_if(is.factor, as.numeric) %>% 
  -1

all_comorb <- all_comorb %>% 
  mutate(total_comorb_by_patient = c(rowSums(all_comorb)), 
         COVIDseverity = d[,73], age_at_encounter = d[,11], service_date = as.Date(d[,9]))

all_comorb %>% 
  summarize(mean_total = mean(total_comorb_by_patient), med_total = median(total_comorb_by_patient), max_total = max(total_comorb_by_patient),
            var_total = var(total_comorb_by_patient))
detach(d)

attach(all_comorb)

colSums(all_comorb[,1:21])

grouped_comorb <- data.frame( 
  select(all_comorb, contains("J")) %>% 
    sum(), 
  select(all_comorb, contains("K", ignore.case = FALSE)) %>% 
    sum(), 
  select(all_comorb, contains("E", ignore.case = FALSE), contains("nutrition")) %>% 
    sum(), 
  select(all_comorb, contains("I", ignore.case = FALSE) & 
         !contains("E", ignore.case = FALSE) & 
         !contains("K", ignore.case =  FALSE) &
         !contains("COVIDseverity")) %>% 
    sum(),
  sum(comorb_chronic_kidney_disease_N18), sum(comorb_hemoptysisR042),                  sum(comorb_lung_transplant_Z942), 
  sum(comorb_nicotine_dependence_F17)) %>% 
    colSums() %>% 
  melt(value.name = 'Patients') %>% 
  mutate(comorbidity = c("Respiratory", "GI", "Nutrition", 
                  "Cardiovascular", "Chronic Kidney Disease", "Hemoptysis", "Lung Transplant", "Nicotine Dependence"))



grouped_comorb %>% 
  ggplot(aes(x = comorbidity, y = Patients)) +
  geom_bar(stat = 'identity')

all_comorb %>% 
  ggplot(aes(x = total_comorb_by_patient)) +
  geom_bar(data = subset(all_comorb, total_comorb_by_patient > 0), fill = "lightsteelblue") +
  scale_x_continuous(breaks = c(1:10)) +
  xlab("Total Number of Comorbidities") +
  ylab("Patients")

all_comorb %>% 
  ggplot(aes(x = total_comorb_by_patient)) +
  geom_bar(fill = "lightsteelblue") +
  scale_x_continuous(breaks = c(0:10)) +
  xlab("Total Number of Comorbidities per Patient") +
  ylab("Patients") +
  facet_wrap(COVIDseverity)


```


```{r}
all_comorb %>% 
  ggplot(aes(x = total_comorb_by_patient, fill = COVIDseverity)) +
  geom_bar() +
  scale_x_continuous(breaks = c(0:10)) +
  xlab("Total Number of Comorbidities per Patient") +
  ylab("Patients") 
 


```


```{r}
all_comorb %>% 
  ggplot(aes(x = COVIDseverity)) +
  geom_bar(fill = "lightsteelblue") +
  xlab("COVID_19 Severity") +
  ylab("Patients") +
  facet_wrap(total_comorb_by_patient)
```


```{r}
include <- ggplot(all_comorb, aes(total_comorb_by_patient, 
                                  fill = as.factor(age_at_encounter))) +
  geom_bar(show.legend = FALSE) +
  scale_x_continuous(breaks = c(0:10)) +
  labs(title = 'Age: {as.integer(frame_time)}', x = 'Total Comorbidities per Patient',
       y = 'Number of Patients') +
  transition_time(age_at_encounter) +
  ease_aes('linear') +
  theme_bw() +
  theme(text = element_text(size = 15)) +
  scale_fill_manual(values = c(viridis_pal(option = "C")(26)))
animate(include, fps = 4)


```


```{r}
ggplot(all_comorb, aes(total_comorb_by_patient, fill = COVIDseverity)) +
  geom_bar(position = "fill") +
  scale_x_continuous(breaks = c(0:10)) +
  labs(x = 'Comorbidities per Patient', y = 'Proportion',
       fill = 'COVID-19 Severity') +
  scale_fill_manual(values = c('#EDD9A3', '#F2637F', '#872CA2'), labels = 
                      c("Mild", "Moderate", "Severe")) +
  theme_bw() +
  theme(text = element_text(size = 15))
```


```{r}
age_sever <- ggplot(all_comorb, aes(age_at_encounter, fill = COVIDseverity)) +
  geom_bar(position = "fill", show.legend = FALSE) +
  scale_x_continuous(breaks = c(0:25)) +
  labs(x = 'Age', y = 'Proportion',
       fill = 'COVID-19 Severity') +
  scale_fill_manual(values = c('#EDD9A3', '#F2637F', '#872CA2'), labels = 
                      c("Mild", "Moderate", "Severe")) +
  theme_bw() +
  theme(text = element_text(size = 15))
```



```{r}
weirdograph <- 
  all_comorb %>% 
  filter(COVIDseverity > 0) %>% 
  ggplot(aes(total_comorb_by_patient,
         fill = COVIDseverity)) +
  geom_bar(position = "fill", aes(color = COVIDseverity)) +
  scale_x_continuous(breaks = c(0:10)) +
  labs(title = 'Age: {as.integer(frame_time)}', x = 'Total Comorbidities per Patient',
       y = 'Number of Patients') +
  transition_time(age_at_encounter) +
  theme_bw() +
  theme(text = element_text(size = 15))

animate(weirdograph, nframes = 25, fps = 1)

  
```

```{r}
 weirdo2 <-  all_comorb %>% 
  filter(COVIDseverity > 0) %>% 
  ggplot(aes(total_comorb_by_patient,
         fill = COVIDseverity)) +
  geom_bar() +
  scale_x_continuous(breaks = c(0:10)) +
  labs(title = 'Age: {as.integer(frame_time)}', x = 'Total Comorbidities per Patient',
       y = 'Number of Patients') +
  transition_time(age_at_encounter) +
  theme_bw() +
  theme(text = element_text(size = 15))
weirdo2 + scale_color_brewer()
  animate(weirdo2, nframes = 25, fps = 1)
```


```{r}
all_comorb %>% 
  ggplot(aes(age_at_encounter)) +
  geom_density(aes(fill = factor(total_comorb_by_patient))) +
  facet_wrap(COVIDseverity)
```


```{r}
all_comorb <- all_comorb %>% 
  mutate(resp_total = as.numeric(rowSums(select(all_comorb, contains("J")))), gi_total =
           as.numeric(rowSums(select(all_comorb, contains("K", ignore.case = FALSE)))), other_total = as.numeric(rowSums(select(all_comorb, contains("comorb") &                                                     !contains("J") & !contains("K", ignore.case = FALSE)                                 & !contains("patient")))))

all_comorb_clean <- all_comorb[,c(1:21)] %>% 
  mutate_if(is.numeric, as.factor) %>% 
  cbind(all_comorb[,c(22:28)]) %>% 
  select(-service_date)
 

col_stackt_df <- 
  rbind(
  all_comorb %>% 
  filter(resp_total > 0 & gi_total == 0 & other_total == 0) %>%
  select(COVIDseverity, age_at_encounter) %>% 
  mutate(class = 'Respiratory'), 
  
  all_comorb %>% 
  filter(resp_total == 0 & gi_total > 0 & other_total == 0) %>% 
  select(COVIDseverity, age_at_encounter) %>% 
  mutate(class = 'GI'),
  
  all_comorb %>% 
  filter(resp_total == 0 & gi_total == 0 & other_total > 0) %>% 
  select(COVIDseverity, age_at_encounter) %>% 
  mutate(class = 'Other'), 
  
  all_comorb %>% 
  filter(resp_total == 0 & gi_total == 0 & other_total == 0) %>%
  select(COVIDseverity, age_at_encounter) %>% 
  mutate(class = 'None'))
```


```{r}
col_stackt_df %>% 
  ggplot(aes(x = COVIDseverity, fill = class)) +
  geom_bar(position = 'dodge')
```


```{r}
percent_sever <- col_stackt_df %>%
    count(COVIDseverity, class) %>%
    group_by(COVIDseverity) %>%
    mutate(percent = (n / sum(n)) * 100) %>%
    ungroup()

percent_sever_no <- col_stackt_df %>%
    filter(class != "None") %>% 
    count(COVIDseverity, class) %>%
    group_by(COVIDseverity) %>%
    mutate(percent = (n / sum(n)) * 100) %>%
    ungroup()

```

```{r}
percent_sever %>% 
  ggplot(aes(x = COVIDseverity, y = percent, fill = class)) +
  geom_bar(stat = 'identity', position = 'dodge')



percent_sever_no %>% 
  ggplot(aes(x = COVIDseverity, y = percent, fill = class)) +
  geom_bar(stat = 'identity', position = 'dodge')
```


```{r}
col_stackt_df %>% 
  ggplot(aes(x = age_at_encounter, after_stat(density), color = class)) +
  geom_freqpoly(binwidth = 5)
```


Preparing data frame for LASSO, I am classifying all cases that require oxygen
intervention, identified by "1" and "2" under the COVIDseverity variable in the
original data set, as severe, which will now be identified as "1". All mild
cases will remain as "0".
```{r}
lasso_comorb <- all_comorb_clean[,c(1:24)] %>% 
  mutate(COVIDseverity = case_when(COVIDseverity > "0" ~ "1",
                                   COVIDseverity == "0" ~ "0")) %>% 
  mutate(COVIDseverity = as.factor(COVIDseverity))

```


LASSO ?!!!?!!!
```{r}
set.seed(36323)
x <-  model.matrix(COVIDseverity ~., lasso_comorb)[, -1]
y <-  lasso_comorb$COVIDseverity
dim(x)
grid <- 10^seq(10, -2, length = 100)

train <-  sample(nrow(x), size = floor(.75 * nrow(x)))
test <-  (-train)
y_test <-  y[test]

lasso_mod <-  glmnet(x[train, ], y[train], alpha = 1,
                     lambda = grid, family = "binomial")

cv_out <- cv.glmnet(x[train, ], y[train], alpha = 1, family = "binomial")

best_lam <-  cv_out$lambda.min

lasso_pred <-  predict(lasso_mod, s = best_lam, newx = x[test, ])

out <-  glmnet(x, y, alpha = 1, lambda = grid, family = "binomial")

lasso_coef = predict(out, type = "coefficients",
                         s = best_lam)[1:24, ]

lasso_coef[lasso_coef != 0]
```


How 'bout linear discriminant analysis?????!!!!!!!!!?!?!?!?!?
```{r}
library(MASS)

lda_comorb <- all_comorb_clean[, c(1, 6:8, 14, 16:19, 21:23)]

train_lda <-  sample(nrow(lda_comorb), size = floor(.75 * nrow(lda_comorb)))

train_lda.df <-  as.data.frame(lda_comorb[train_lda,])

test_lda.df <-  as.data.frame(lda_comorb[-train_lda,])

```

lord help me

```{r}
f <- paste(names(train_lda.df)[12], "~", paste(names(train_lda.df)[-12],
                                               collapse = " + "))

lda_fit <- lda(as.formula(paste(f)), data = train_lda.df)
lda_fit

plot(lda_fit)

lda_class <- predict(lda_fit, test_lda.df)$class
```


```{r}
table(lda_class, test_lda.df[,"COVIDseverity"])
```


```{r}
test_cs <- test_lda.df[,"COVIDseverity"]
lda_class <- as.ordered(lda_class)
mean(lda_class == test_cs)
```

```{r}
lr_comorb <- glm(COVIDseverity ~ comorb_bronchiectasis_J47 + comorb_pneumothorax_J93 + comorb_resp_failure_J96 + comorb_malnutrition + comorb_other_GI_notLiver_K_excludesK70K77 + comorb_hypertensive_heart_disease_I11 + comorb_chronic_kidney_disease_N18 + comorb_nicotine_dependence_F17 + comorb_heart_failure_I50 + comorb_lung_transplant_Z942 + total_comorb_by_patient, family = "binomial", data = lasso_comorb)

summary(lr_comorb)
```


```{r}
library(heplots)
boxM(lda_comorb[,1:11], lda_comorb[,12])
```

Sooooooooooo we don't meet the assumptions necessary for LDA :/


```{r}
all_comorb_clean <-  all_comorb_clean %>%
  mutate(
  age_group = case_when(age_at_encounter <= 12 ~ "1. child",
                        age_at_encounter > 12 & age_at_encounter <= 17 ~                                "2. teen",
                        age_at_encounter > 17 ~ "3. adult") ) %>% 
  mutate_if(is.character, as.factor)

con_comorb <- all_comorb_clean[, c("COVIDseverity",
                            "total_comorb_by_patient", "age_group")] %>% 
  mutate(severity = case_when(COVIDseverity > "0" ~ "1",
                                   COVIDseverity == "0" ~ "0")) %>% 
  filter(total_comorb_by_patient <= 1) %>% 
  mutate(severity = as.numeric(severity), comorb = total_comorb_by_patient) %>% 
  as.matrix()


```

```{r}

con_table <- xtabs(~ age_group + comorb + severity, 
                   data = con_comorb)
ftable(prop.table(con_table, c(1,2)))
```


```{r}

ftable(con_table)
```


```{r}
comorb <- c("0", "1")
severity <-  c("mild", "severe")
age <- c("child", "teen", "adult")

table <-  expand.grid(comorb = comorb, severity = severity,
                      age = age)

count <-  c(1123, 674, 76, 59, 1094, 487, 44, 15, 1879, 1116, 164, 79)

table <- cbind(table, count = count)

table
```


```{r}
pls <- xtabs(count ~ comorb + severity + age, table)
pls
```


```{r}
ftable(pls)
```


```{r}
Frequency <- as.vector(margin.table(pls, 3))

CumFrequency <- cumsum(Frequency)

cbind(age,Frequency = Frequency,Percentage=Frequency/sum(Frequency),
      CumFrequency = CumFrequency,
      CumPercentage = CumFrequency/sum(Frequency))
 

 
Frequency <- as.vector(margin.table(pls, 2))
CumFrequency <- cumsum(Frequency)
cbind(severity, Frequency = Frequency,
      Percentage = Frequency/sum(Frequency),
      CumFrequency = CumFrequency,
      CumPercentage = CumFrequency/sum(Frequency))
 

 
Frequency <- as.vector(margin.table(pls,1))

CumFrequency <- cumsum(Frequency)

cbind(comorb,Frequency = Frequency,
      Percentage = Frequency/sum(Frequency),
      CumFrequency = CumFrequency,
      CumPercentage = CumFrequency/sum(Frequency))
```


The following is a test for mutual independence
```{r}
E <- array(NA, dim(pls))
for (i in 1:dim(pls)[1]) {
for (j in 1:dim(pls)[2]) {
for (k in 1:dim(pls)[3]) {
E[i,j,k] = (margin.table(pls,3)[k] * margin.table(pls,2)[j] * margin.table(pls,1))[i] / (sum(pls))^2
}}}

E
```


```{r}
df <- (length(pls) - 1) - (sum(dim(pls)) - 3)

X2 <-  sum((pls - E)^2 / E)
X2

1 - pchisq(X2, df)

G2 <- 2 * sum(pls * log(pls / E))
G2

1 - pchisq(G2, df)
```
The tests imply that the variables for COVID severity, age group, and presence 
of one or no comorbidities are not mutually independent


Test for joint independence of (severity, age group and comorbidity)
```{r}
ji_sac <- ftable(pls, row.vars = c(3, 2))

results <- chisq.test(ji_sac)
results
```
Since the chi-squared value is large and the p-value is close to 0, this indicates that COVID severity is not independent of age group and comorbidity.
This means age_group*comorbidity cannot predict covid severity.

```{r}
mantelhaen.test(pls)

mantelhaen.test(pls, correct = FALSE)
```


```{r}


lasso_group <- all_comorb_clean %>% 
  mutate(COVIDseverity = case_when(COVIDseverity > "0" ~ "1",
                                   COVIDseverity == "0" ~ "0"))

lasso_group <- lasso_group[,c(23, 26:28)]
 

set.seed(36323)
x.2 <-  model.matrix(COVIDseverity ~., lasso_group)[, -1]
y.2 <-  lasso_group$COVIDseverity

grid.2 <- 10^seq(10, -2, length = 100)

train.2 <-  sample(nrow(x.2), size = floor(.75 * nrow(x.2)))
test.2 <-  (-train.2)
y_test.2 <-  y.2[test.2]

lasso_mod.2 <-  glmnet(x.2[train.2, ], y.2[train.2], alpha = 1,
                     lambda = grid.2, family = "binomial")

cv_out.2 <- cv.glmnet(x.2[train.2, ], y.2[train.2], alpha = 1, family = "binomial")

best_lam.2 <-  cv_out.2$lambda.min

lasso_pred.2 <-  predict(lasso_mod.2, s = best_lam.2, newx = x.2[test.2, ])

out.2 <-  glmnet(x.2, y.2, alpha = 1, lambda = grid.2, family = "binomial")

lasso_coef.2 = predict(out.2, type = "coefficients",
                         s = best_lam.2)[1:3, ]

lasso_coef.2[lasso_coef.2 != 0]
```


```{r}
glm_df <- all_comorb_clean %>% 
  mutate(COVIDseverity = as.factor(case_when(COVIDseverity > "0" ~ "Severe",
                                   COVIDseverity == "0" ~ "Mild")))

  
train <-  sample(nrow(all_comorb_clean), size = floor(.75 * nrow(all_comorb_clean)))



glm_fit <- glm(COVIDseverity ~ comorb_bronchiectasis_J47 + 
                 comorb_pneumothorax_J93 + comorb_resp_failure_J96 + 
                 comorb_malnutrition + comorb_nicotine_dependence_F17, 
               data = glm_df, family = "binomial", subset = train)

glm_fit
```


```{r}
glm_probs <- predict(glm_fit, type = "response")
glm_probs[1:10]

contrasts(glm_df[,"COVIDseverity"])
```

```{r}
glm_pred <- rep("Mild", 9337)
glm_pred[glm_probs > .5] = "Severe"

table(glm_pred, glm_fit[,"COVIDseverity"])

mean(glm_pred == glm_fit[,"COVIDseverity"])
```

